---
title: "Compensating and Equivalent Variation"
---

```{r}
source("Utility Maximization.R")
source("Expenditure Minimization.R")
```

@Miles - I've added the functions that came up on my quiz. You're welcome to add in yours - I'll chuck in the others when the need comes up.

Note: CV and EV lower is better

$$ 
CV = e(w^{new}, U^{old}) - M^{new} = e(w^{new}, v(w^{old}, M^{old})) - w^{new}x^{new}\\
EV = M^{old} - e(w^{old}, U^{new}) = w^{old}x^{old} - e(w^{old}, v(w^{new}, M^{new}))
$$

# Cobb Douglas

## 2 Good

### Compensating Variation

```{r}

# Policy changes
policy_changes = list(
  c(1.04, 1.17, 1.06),
  c(1.05, 0.8, 0.92),
  c(0.8, 0.9, 0.95),
  c(1.1, 0.88, 0.94)
)

# Old variables
a1 = 0.2
a2 = 0.8
w1 = 1
w2 = 1
M = 48

# Calculating the old V
v_old = cobb_douglas_two_good_umax(
  1, 
  a1,
  a2,
  w1, 
  w2, 
  M
)

cv = c()

for (policy in policy_changes) {
  # Calculating new expenditure function with new prices and old v
  e_new = cobb_douglas_two_good_emin(
    1, 
    a1,
    a2,
    w1 * policy[1], 
    w2 * policy[2], 
    v_old$v)
  
  # Calculating the compensating variation with new prices and income
  cv = c(cv, e_new$e - M*policy[3])
}

cv
```

### Equivalent Variation

```{r}

```

## 3 Good

### Compensating Variation

```{r}

```

### Equivalent Variation

```{r}

```

# CES

## 2 Good

### Compensating Variation

```{r}

```

### Equivalent Variation

```{r}

```

## 3 Good

### Compensating Variation

```{r}

```

### Equivalent Variation

```{r}
# Policy changes
policy_changes = list(
  c(1.08, 0.87, 1.08, 0.98),
  c(1.08, 1.13, 0.95, 0.95),
  c(0.96, 0.81, 0.87, 0.83),
  c(0.84, 0.81, 0.83, 0.89)
)

# Old variables
a1 = 0.33
a2 = 0.2
a3 = 0.47
w1 = 1
w2 = 1
w3 = 1
sigma = 1.42
M = 44

ev = c()

for (policy in policy_changes) {
  # Calculating the new V
  v_new = ces_three_good_umax(
    1, 
    sigma,
    a1,
    a2,
    a3,
    w1*policy[1], 
    w2*policy[2],
    w3*policy[3],
    M*policy[4]
  )
  
  # Calculating new expenditure function with old prices and new v
  e_new = ces_three_good_emin(
    1, 
    sigma,
    a1,
    a2,
    a3,
    w1, 
    w2, 
    w3, 
    v_new$v
  )
  
  # Calculating the equivalent variation with old income and new utility
  ev = c(ev, M - e_new$e)
}

ev
```

